The Galaxy Clustering Crisis in Abundance Matching

### 2017-05-04 paper/paper.tex
updates

\title[The Galaxy Clustering Crisis]{The Galaxy Clustering Crisis in Abundance Matching}
\author[Campbell et al.]
{\parbox[t]{\textwidth}{
Duncan Campbell$^{1}$\thanks{E-mail: duncan.campbell@yale.edu},
Frank C. van den Bosch$^{1}$,
Nikhil Padmanabhan$^{2, 3}$, \\ 
Andrew Zentner$^{4}$,
Johannes Lange$^{1}$,
Yao-Yuan Mao$^{4}$, and
Fangzhou Jiang$^{5}$
} \\
\vspace*{3pt} \\
$^{1}$Department of Astronomy, Yale University, New Haven, CT 06511, USA\\
$^{2}$Department of Physics, Yale University, New Haven, CT 06520, USA \\
$^{3}$Yale Center for Astronomy and Astrophysics, Yale University, New Haven, CT 06520, USA\\
$^{4}$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260
$^{5}$Racah Institute of Physics, The Hebrew University, Jerusalem 91904, Israel
}

@@

\begin{abstract}
Galaxy clustering on small scales is significantly under-predicted by sub-halo abundance matching (SHAM) models that populate (sub-)haloes with galaxies based on peak halo mass, $M_{\rm peak}$.  SHAM methods based on the peak maximum circular velocity, $V_{\rm peak}$, have had much better success.  The primary reason $M_{\rm peak}$ based models fail is the relatively low abundance of satellite galaxies produced in these models compared to those based on $V_{\rm peak}$.  Despite success in predicting clustering, simple $V_{\rm peak}$ based SHAM models have their own problems, and there is little theoretical reason to prefer $V_{\rm peak}$ over $M_{\rm peak}$ as a proxy for stellar mass.  We identify three possible remedies that could save mass based models.  1.) SHAM models require a significant population of ``orphan'' galaxies as a result of artificial disruption of sub-haloes in modern dark matter simulations, 2.) satellites must grow significantly after the time of accretion, and/or 3.) stellar mass is significantly affected by halo assembly history.  Regardless of the particulars, we show that popular SHAM models based on $M_{\rm peak}$ cannot be correct as presented.  Either $V_{\rm peak}$ truly is a better predictor of stellar mass, or SHAM models are missing vital component(s) that significantly affect galaxy clustering.
\end{abstract}

@@

\begin{figure}
    \includegraphics{figures/f_sat_mstar.pdf}
    \caption{The satellite fraction as a function of stellar mass is plotted for SHAM model as lines.  The models are RM (solid cyan with shaded region showing $\pm 1\sigma$ error on the model prediction), RV (solid magenta), M13 (solid black), Y12 (dashed), and B13 (dotted).  For comparison, we show the satellite fraction as inferred by \citet{vandenBosch:2008fv} using a galaxy group catalogue of SDSS galaxies.}
    \label{fig:f_sat_mstar}
\end{figure}

@@

The projected correlation functions for each model calculated in three stellar mass bins as described above are shown in Fig. \ref{fig:wp_comparison}.  For comparison, we show the corresponding measurement made on a galaxy sample in SDSS \citep[``Mass-limit'' sample in][]{Yang:2012ew}.  The difference between the clustering signal predictions of each SHAM model and observations is striking.  Each halo mass based models\footnote{As a short-hand, we will frequently refer to the RM, M13, B13, and Y12 models as ``mass based models''.} severely under-predicts the clustering signal in the two least massive bins (left two panels), while the RV model, based on $V_{\rm peak}$, if anything, slightly over-predicts the signal.  On the other hand, all the models considered here roughly predict an accurate clustering signal consistent with observations in the most massive stellar mass bin (right panel).

@@

The success of the RV model is consistent with findings by \cite{Reddick:2013gi} who show that SHAM based on $V_{\rm peak}$ most closely matches galaxy clustering observations in SDSS, while all other halo properties under-predict the signal.  We confirm that result here; furthermore, we show that more complicated evolving SMHM models based on $M_{\rm peak}$ do not alleviate the problem.  In fact, evolution seems to exacerbate the clustering decrement of $M_{\rm peak}$ based SHAM as evidenced by the RM model producing the strongest clustering signal amongst the mass based models.  The remainder of this paper examines why these models fail to produce strong enough clustering, why the RV model (and similar models) succeeds, and possible solutions to ``save'' mass based SHAM models (under the assumption these models are worth saving).

@@

In this section, we examine the origin of the success of $V_{\rm peak}$ based SHAM models.  By comparing the differences in $V_{\rm peak}$ and $M_{\rm peak}$ based models, we identify the culprit(s) in the failure of the mass based models to match observational clustering measurements.  We also discuss some reasons to prefer mass based models over $V_{\rm peak}$ based models.

@@

Where clustering is under-predicted, the problem is most severe at small scales $(<1~h^{-1}{\rm Mpc})$.  This is a strong indication that the culprit is satellite galaxies, or more precisely, a lack of satellite galaxies in the mass based SHAM mocks.  With this in mind, we examine the satellite fraction, $f_{\rm sat}$, in each mock in Fig. \ref{fig:f_sat_mstar}.  The RV model has a larger satellite fraction, $\sim 27\%$, relative to the other models which are below $\sim 22\%$.  Comparing the clustering signals in Fig. \ref{fig:wp_comparison} with the satellite fractions in Fig. \ref{fig:f_sat_mstar} shows that there is a nearly one-to-one correspondence between the clustering strength on small scales and $f_{\rm sat}$ of each model.  Furthermore, the satellite fractions between the models do not diverge until approximately $5 \times 10^{10}h^{-2}M_{\odot}$, above which the models largely agree with the clustering results.  Considering this, it may be more appropriate to restate the under-prediction of galaxy clustering in these models as an underproduction of satellite galaxies.

@@

It is the correlated scatter in $V_{\rm peak}$ at fixed $M_{\rm peak}$ which is responsible for the difference in the satellite fractions between the RV and RM models.  By populating (sub-)haloes by their rank on $V_{\rm peak}$, more sub-haloes will be populated than would have been had $M_{\rm peak}$ instead been utilized for any given stellar mass threshold sample.  This same reasoning carries over to the other mass based models.

@@

Given that the RV and mass based models make different assumptions for the SMHM relation, it is not surprising that each model predicts different galaxy clustering signals and satellite abundances.  Up to this point, one might be inclined to favour the RV model over the mass based SHAM models.  Here we show that the RV model implies galaxy growth histories that are at odds with the other models and observations.

We have shown that in the RV model the stellar mass fraction relative to peak halo mass is larger for satellites than centrals.  This implies that the SMHM relation decreases over time (moves towards the right in Fig. \ref{fig:SMHM_comparison}) in the RV model while it increases in the other models (moves towards the left).  This conclusion rests on two assumptions of the RV model: satellites do not evolve significantly in mass after accretion, and the $M_* - V_{\rm peak}$ relation itself is static (an implicit assumption of the RV model).  We examine the second assumption about the growth of stellar mass in centrals, and leave an examination of post-accretion satellite mass growth to \S \ref{sec:sat_growth}.

\begin{figure}
    \includegraphics{figures/mstar_of_z_rm.pdf}
    \caption{The average stellar mass as a function of redshift divided by the mass at z=0 for haloes of various values of $V_{\rm peak}$ at z=0 (coloured lines).  The prediction from the non-evolving RV model is shown as solid lines.  The prediction of the evolving M13 model is shown as dotted lines.  $V_{\rm peak}(z=0)$ was converted $M_{\rm vir}$ using a fitting function in order to directly compare the RV and M13 models.  The implied average stellar mass growth histories of haloes are drastically different between the two models.  The RV model predicts {\rm much} earlier stellar mass growth relative to the M13 model. {\bf [DC: fix color table]}}
    \label{fig:mstar_of_z_rm}
\end{figure}

@@

To begin, we fit the $\langle M_* \rangle - V_{\rm peak}$ (SMVP) relation in the RV model with a function of the form:
%
\begin{linenomath}
\begin{equation}
\label{eq:smvp_relation}
\langle M_* | V_{\rm peak} \rangle = 2M_0\left(\frac{V_{\rm peak}}{V_0}\right)\left[\left(\frac{V_{\rm peak}}{V_0}\right)^{\alpha}+\left(\frac{V_{\rm peak}}{V_0}\right)^{\beta}\right]^{-1}
\end{equation}
\end{linenomath}
%
We perform a non-linear least squares fit to the mean stellar mass in 0.025 dex $V_{\rm peak}$ bins.  We find the parameters $\log(M_0/h^{-2}M_{\odot})=10.0 \pm 0.01$, $\log(V_0/{\rm km/s})=2.197 \pm 0.004$, $\alpha=-4.6 \pm 0.1$, and $\beta=-0.20 \pm 0.01$ provide an excellent fit.

As discussed, the RV model implies that galaxies grow along a fixed SMVP, i.e. galaxy growth is tied to the growth of $V_{\rm peak}$ in haloes.  Accordingly, the average growth history of galaxies is dependent on the average potential well growth histories of haloes, $\langle V_{\rm peak} \rangle (z | V_{\rm peak, 0})$, \citep[PWGHs,][]{Bosch:2014cu}, and the form of the SMVP relation, eq. \ref{eq:smvp_relation}.  We show the implied growth history of galaxies in the RV model in Fig. \ref{fig:mstar_of_z_rm} for haloes with different $V_{\rm peak}$ at z=0.0.  For comparison, we also show the stellar mass growth histories from M13, which depends on the average mass accretion histories (MAHs) of haloes, where halo mass has been converted to $V_{\rm peak}$ using eq. \ref{eq:vpeak} to place these histories on the same figure.  B13 and Y12 result in qualitatively similar growth histories as M13.

The stellar mass growth histories predicted by the two models are drastically different.  Because the potential wells of haloes (and as a result $V_{\rm peak}$) are set early, galaxies in the RV model grow rapidly at early times and more slowly at low redshift.  This is phenomenologically at odds with the predictions of the evolving SHAM models in which galaxies grow much more slowly at high redshift and more rapidly at later times.  The effect of this can be seen in the evolution of the stellar mass function as shown in Fig. \ref{fig:stellar_mass_function_evolution}.  The RV model over-produces galaxies too early, while the M13 model was tuned to reproduce $\phi(M_*,z)$.  The Y12 and B13 models predict broadly consistent growth histories and $\phi(M_*,z)$ as M13.

@@

Given that the M13, Y12, and B13 models were fit to the stellar mass function at various redshifts and reproduce the cosmic star-formation density in the Universe, a model with such radically different stellar mass growth histories like RV is unlikely to be consistent with these same measurements.  This is one reason a simple model based on $V_{\rm peak}$ like RV is likely deeply flawed.  However, it is also likely that a more complicated $V_{\rm peak}$ based SHAM model could produce realistic growth histories, but this would require evolution in the $M_* - V_{\rm peak}$ relation and likely a modification to the evolution of satellite galaxies.  Given the necessity of these modifications, it is not clear that there is a reason to prefer a $V_{\rm peak}$ based model.  Beyond this Occam's razor-like argument, there are other problems with $V_{\rm peak}$ driving the stellar mass mapping into haloes.  For any given (sub-)halo, $V_{\rm peak}$ is generally set during major mergers \citep[1:5 or larger, ][]{Behroozi:2014bg}, and if stellar mass is tied to $V_{\rm peak}$, this could imply very bursty star-formation closely tied to mergers, and there is little evidence this is the case {\bf [citations needed]}.  On the other-hand, there is theoretical reasoning to prefer halo mass as driving the mapping.  The peak halo mass is a good indicator of the amount of baryons that have been available to a galaxy to grow over its history {\bf [citations needed]}, assuming that this quantity is proportional to the baryon fraction, $f_{\rm bary}$.

Our comparison between mass based models and our $V_{\rm peak}$ based models makes it clear that there is a tension in current SHAM models between fitting clustering while simultaneously reproducing the growth of stellar mass in the Universe.  This presents a ``clustering'' crisis for SHAM.  Our exploration of why $V_{\rm peak}$ succeeds, and the differences compared to the mass based models suggests multiple ways abundance matching could be altered in order to alleviate this crisis.  Given that an obvious difference between the RV model and the others is the increased satellite fraction, one solution is to simply increase the number of satellites in the mass based models.  In the following section, we examine the plausibility of missing sub-haloes, and thus satellites, in our implementation of the mass based SHAM models.  A significant population of missing sub-haloes would be an indication that so called ``oprhan'' galaxies play an important role.  Apart from orphans, we also consider two other scenarios to increase the satellite fraction--post-accretion satellite growth in \S \ref{sec:sat_growth} and assembly bias effects in \S \ref{sec:assem_bias}.

@@

Finally, we ask ``how many orphans are needed to increase the clustering signal sufficiently in the SHAM models based on $M_{\rm peak}$?''  To answer this question, we adjust $C_0$ in eq. (\ref{eq:sub_completeness}), while keeping the other parameters fixed, and fit to the galaxy clustering observations.  Lower values of $C_0$ imply an overall increased population of clone sub-haloes available to host orphans at all masses.  This correction assumes that sub-haloes are being artificially disrupted or merged with the host {\em at all masses}.  For each model, $C_0$ was adjusted in order to best fit the galaxy clustering signal in the intermediate mass bin $(10.0 \leq \log(M_*/h^{-2}M_{\odot}) <10.5)$.  The result of this exercise is shown in Fig. \ref{fig:wp_comparison_w_orphans} using the {\tt sub-profile} method to assign the positions and velocities.  We find that each model requires $C_{0} = [0.6,1.0, 0.4, 0.5, 0.5]$, for the RM, RV, M13, Y12, and B13 models respectively.  While the RV model does not require orphans, the mass based SHAM models require that on average approximately {\em half of all satellites are orphans}.  In Fig. \ref{fig:orphan_clustering_2} we show how assigning orphans positions within their host affect the clustering predictions for the M13 model in the intermediate mass bin only.  As expected, using {\rm dm-profile} boosts the clustering signal at small scales relative to {\rm sub-profile}, but the effect is small compared to the dependence on $C_{0}$.  This trend is largely similar at the other mass considered and for the other SHAM models.

We make no attempt to fit for new parameters in the evolving SMHM models using our orphan model and clustering measurements, while the RM and RV models adjust automatically to the increased abundance of sub-haloes.  Adjusting the population of sub-haloes so drastically in the evolving models will have an effect on the parameter inference for the SMHM relation.  With this caveat in mind, the most noticeable failure of the mass based models is the over-prediction of the small scale ($< 1 {\rm Mpc}$) clustering signal in the most massive stellar mass bin ($10.5\leq \log(M_*/h^{-2}M_{\odot})<11.0$).  Each of these models now produce too many massive satellites.  This problem could be reduced by altering the SMHM for massive satellites or reducing the number of massive orphans.  We leave a detailed study on the self-consistently including orphan galaxies in SHAM to future work.

@@

While we have shown that a large population of ``orphan'' galaxies would alleviate the clustering crisis in mass based SHAM models, high resolution dark matter simulations do not seem to provide evidence of the requisite missing sub-halo population.  With this in mind, we explore alternative mechanisms to boost the galaxy clustering signal in SHAM models.  In this section, we consider whether continued star-formation in satellites after accretion can significantly boost clustering.  This is motivated by findings that suggest satellites continue to form stars and grow in stellar mass after accretion for between $\sim 2$ and $4$ Gyr \citep{Wetzel:2013dw}.  If satellites continue to grow after accretion, SHAM methods which use the SMHM relation at the time of accretion to assign stellar mass to satellites will under-estimate their stellar mass (e.g. M13 and B13).

@@

In this section we examine whether introducing assembly bias into the galaxy-halo mapping can increase the satellite fraction and galaxy clustering signal in mass based SHAM models.  Specifically, we consider a model where $M_{*}$ is correlated with the formation history of the (sub-)halo it occupies such that early forming haloes host more massive galaxies than late forming haloes at {\em fixed} peak (sub-)halo mass.  While many measures of halo formation history have been employed, for this work we use the redshift a halo first achieves a fraction, $\rm f$, of its peak halo mass, $z_{\rm f}$ (see Appendix \ref{sec:halo_properties} for details on how $z_{\rm f}$ is calculated).

@@

To induce a correlation between $M_{*}$ and $z_{\rm f}$ we bin (sub-)haloes in small 0.1 dex $M_{\rm peak}$ bins.  We then rank order (sub-)haloes by $z_{\rm f}$ and galaxies by $M_{*}$, re-assigning the most massive galaxies to the earliest forming (sub-)haloes in the bin.  We use ${\rm f}=0.5$, the redshift where a (sub-)halo reaches half its peak mass.  Furthermore, we parametrize the strength of this correlation by the Spearman’s rank order correlation coefficient, $\rho(M_{\rm vir})$.  We weaken the correlation by degrading the rank ordering as described in appendix \ref{appendix:ranks} in order to reduce the effect of assembly bias on $M_*$.

>>>>

\title[The Galaxy Clustering Crisis]{The Galaxy Clustering Crisis in Abundance Matching}
\author[Campbell et al.]
{\parbox[t]{\textwidth}{
Duncan Campbell$^{1}$\thanks{E-mail: duncan.campbell@yale.edu},
Frank C. van den Bosch$^{1}$,
Nikhil Padmanabhan$^{2, 3}$, \\ 
Andrew Zentner$^{4}$,
Johannes Lange$^{1}$,
Yao-Yuan Mao$^{4}$, and
Fangzhou Jiang$^{5}$
} \\
\vspace*{3pt} \\
$^{1}$Department of Astronomy, Yale University, New Haven, CT 06511, USA\\
$^{2}$Department of Physics, Yale University, New Haven, CT 06520, USA \\
$^{3}$Yale Center for Astronomy and Astrophysics, Yale University, New Haven, CT 06520, USA\\
$^{4}$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260\\
$^{5}$Racah Institute of Physics, The Hebrew University, Jerusalem 91904, Israel
}

@@

\begin{abstract}
Galaxy clustering on small scales is significantly under-predicted by sub-halo abundance matching (SHAM) models that populate (sub-)haloes with galaxies based on peak halo mass, $M_{\rm peak}$.  SHAM methods based on the peak maximum circular velocity, $V_{\rm peak}$, have had much better success.  The primary reason $M_{\rm peak}$ based models fail is the relatively low abundance of satellite galaxies produced in these models compared to those based on $V_{\rm peak}$.  Despite success in predicting clustering, simple $V_{\rm peak}$ based SHAM models have their own problems, and there is little theoretical reason to prefer $V_{\rm peak}$ over $M_{\rm peak}$ as a proxy for stellar mass.  We identify three possible remedies that could save mass-based models.  1.) SHAM models require a significant population of ``orphan'' galaxies as a result of artificial disruption of sub-haloes in modern dark matter simulations, 2.) satellites must grow significantly after the time of accretion, and/or 3.) stellar mass is significantly affected by halo assembly history.  Regardless of the particulars, we show that popular SHAM models based on $M_{\rm peak}$ cannot be correct as presented.  Either $V_{\rm peak}$ truly is a better predictor of stellar mass, or SHAM models are missing vital component(s) that significantly affect galaxy clustering.
\end{abstract}

@@

\begin{figure}
    \includegraphics{figures/f_sat_mstar.pdf}
    \caption{The satellite fraction as a function of stellar mass is plotted for SHAM model as lines.  The models are RM (solid cyan), RV (solid magenta with the shaded region showing $\pm 1\sigma$ error on the model prediction), M13 (solid black), Y12 (dashed), and B13 (dotted).  For comparison, we show the satellite fraction as inferred by \citet{vandenBosch:2008fv} using a galaxy group catalogue of SDSS galaxies.}
    \label{fig:f_sat_mstar}
\end{figure}

@@

The projected correlation functions for each model calculated in three stellar mass bins as described above are shown in Fig. \ref{fig:wp_comparison}.  For comparison, we show the corresponding measurement made on a galaxy sample in SDSS \citep[``Mass-limit'' sample in][]{Yang:2012ew}.  The difference between the clustering signal predictions of each SHAM model and observations is striking.  Each halo mass-based models\footnote{As a short-hand, we will frequently refer to the RM, M13, B13, and Y12 models as ``mass-based models''.} severely under-predicts the clustering signal in the two least massive bins (left two panels), while the RV model, based on $V_{\rm peak}$, if anything, slightly over-predicts the signal.  On the other hand, all the models considered here roughly predict an accurate clustering signal consistent with observations in the most massive stellar mass bin (right panel).

@@

The success of the RV model is consistent with findings by \cite{Reddick:2013gi} who show that SHAM based on $V_{\rm peak}$ most closely matches galaxy clustering observations in SDSS, while all other halo properties under-predict the signal.  We confirm that result here; furthermore, we show that more complicated evolving SMHM models based on $M_{\rm peak}$ do not alleviate the problem.  In fact, evolution seems to exacerbate the clustering decrement of $M_{\rm peak}$ based SHAM as evidenced by the RM model producing the strongest clustering signal amongst the mass-based models.  The remainder of this paper examines why these models fail to produce strong enough clustering, why the RV model (and similar models) succeeds, and possible solutions to ``save'' mass-based SHAM models (under the assumption these models are worth saving).

@@

In this section, we examine the origin of the success of $V_{\rm peak}$ based SHAM models.  By comparing the differences in $V_{\rm peak}$ and $M_{\rm peak}$ based models, we identify the culprit(s) in the failure of the mass-based models to match observational clustering measurements.  We also discuss some reasons to prefer mass-based models over $V_{\rm peak}$ based models.

@@

Where clustering is under-predicted, the problem is most severe at small scales $(<1~h^{-1}{\rm Mpc})$.  This is a strong indication that the culprit is satellite galaxies, or more precisely, a lack of satellite galaxies in the mass-based SHAM mocks.  With this in mind, we examine the satellite fraction, $f_{\rm sat}$, in each mock in Fig. \ref{fig:f_sat_mstar}.  The RV model has a larger satellite fraction, $\sim 27\%$, relative to the other models which are below $\sim 22\%$.  Comparing the clustering signals in Fig. \ref{fig:wp_comparison} with the satellite fractions in Fig. \ref{fig:f_sat_mstar} shows that there is a nearly one-to-one correspondence between the clustering strength on small scales and $f_{\rm sat}$ of each model.  Furthermore, the satellite fractions between the models do not diverge until approximately $5 \times 10^{10}h^{-2}M_{\odot}$, above which the models largely agree with the clustering results.  Considering this, it may be more appropriate to restate the under-prediction of galaxy clustering in these models as an underproduction of satellite galaxies.

@@

It is the correlated scatter in $V_{\rm peak}$ at fixed $M_{\rm peak}$ which is responsible for the difference in the satellite fractions between the RV and RM models.  By populating (sub-)haloes by their rank on $V_{\rm peak}$, more sub-haloes will be populated than would have been had $M_{\rm peak}$ instead been utilized for any given stellar mass threshold sample.  This same reasoning carries over to the other mass-based models.

@@

\begin{figure}
    \includegraphics{figures/mstar_of_z_rm.pdf}
    \caption{The average stellar mass as a function of redshift divided by the mass at z=0 for haloes of various values of $V_{\rm peak}$ at z=0 (coloured lines).  The prediction from the non-evolving RV model is shown as solid lines.  The prediction of the evolving M13 model is shown as dotted lines.  $V_{\rm peak}(z=0)$ was converted $M_{\rm vir}$ using a fitting function in order to directly compare the RV and M13 models.  The implied average stellar mass growth histories of haloes are drastically different between the two models.  The RV model predicts {\rm much} earlier stellar mass growth relative to the M13 model.}
    \label{fig:mstar_of_z_rm}
\end{figure}

\begin{figure}
    \includegraphics{figures/stellar_mass_zform.pdf}
    \caption{The median redshift at which a central galaxy reaches a fraction, $f$, of its $z=0$ stellar mass as a function $V_{\rm peak}$ at $z=0$. The three panels are for $f=0.9, 0.5$, and $0.1$ from top to bottom, respectively.  The lines are for the RM (cyan), RV (solid magenta), M13 (solid black), Y12 (dashed), and B13 (dotted).  The RV model consistently forms a larger fraction of galaxies' mass at higher redshift relative to the mass-based models.  The only exception to this rule is the Y12 model, where massive galaxies loose mass at low redshift.}
    \label{fig:mstar_zform}
\end{figure}

Given that the RV and mass-based models make different assumptions for the SMHM relation, especially regarding satellites, it is not be surprising that each model predicts different galaxy clustering signals and satellite abundances.   Considering the RV model's success in fitting clustering observations and the mass-based models failure (see \S \ref{sec:clustering}), it appears that the RV model should be favoured as the more physical.  However, in this section we show that the RV model implies galaxy growth histories that are incompatible with both the other models and observations.

% Up to this point, one might be inclined to favour the RV model over the mass-based SHAM models given its success in fitting galaxy clustering observations.

In the previous section we showed that, in the RV model, the stellar masses of satellites are on average larger than centrals at fixed $M_{\rm peak}$.  This finding implies that the SMHM relation in the RV model decreases over time (evolves towards the right in Fig. \ref{fig:SMHM_comparison}), while it increases in the other models (evolves from right to left).  This conclusion rests on two implicit assumptions of the RV model: the stellar mass of satellites does not evolve significantly after $z_{\rm acc}$, and the $M_* - V_{\rm peak}$ (SMVP) relation itself does not evolve.  A non-evolving SMVP implies that galaxies grow along the SMVP, i.e. galaxy growth is tied to the growth of $V_{\rm peak}$.  In this section we examine the latter assumption, and we leave an examination of post-accretion satellite evolution to \S \ref{sec:sat_growth}.

@@

In order to extrapolate the SMVP relation in the RV model to less massive haloes, we fit it with a function of the form:
%
\begin{linenomath}
\begin{equation}
\label{eq:smvp_relation}
\langle M_* | V_{\rm peak} \rangle = 2M_0\left(\frac{V_{\rm peak}}{V_0}\right)\left[\left(\frac{V_{\rm peak}}{V_0}\right)^{\alpha}+\left(\frac{V_{\rm peak}}{V_0}\right)^{\beta}\right]^{-1}
\end{equation}
\end{linenomath}
%
We perform a non-linear least squares fit to the median stellar mass in 0.025 dex $V_{\rm peak}$ bins.  We find the parameters $\log(M_0/h^{-2}M_{\odot})=10.0 \pm 0.01$, $\log(V_0/{\rm km/s})=2.197 \pm 0.004$, $\alpha=-4.6 \pm 0.1$, and $\beta=-0.20 \pm 0.01$ provide an excellent fit.  We perform a similar fit to the SMHM relation in the RM model, assuming it does not evolve.

In addition to the form of the SMVP relation, the average growth history of galaxies in RV is dependent on the potential well growth history (PWGH) of haloes, $\langle V_{\rm peak}| V_{\rm peak}(z=0) \rangle(z)$.  Using the PWGHs from \citet{Bosch:2014cu}, in Fig. \ref{fig:mstar_of_z_rm} we show the implied growth history of galaxies in the RV model for host-haloes with different $V_{\rm peak}$ at z=0.  For comparison, we also show stellar mass growth histories from M13, which instead depend on the average mass accretion histories (MAHs) of haloes and the explicitly parametrized evolution of the SMHM relation.  Note that we have converted $M_{\rm peak}(z=0)$ to $V_{\rm peak}(z=0)$ using eq. \ref{eq:vpeak} to place these histories on the same figure.

Fig. \ref{fig:mstar_of_z_rm} shows that, in the RV model, galaxy growth is largely self-similar--at any given redshift, galaxies of every mass are growing at similar rates.  Conversely, in the M13 model, galaxy growth is much more dependent on halo mass--high mass galaxies form early, growing slowly at low redshift, while low mass galaxies form late.  Slow late time growth of massive galaxies is necessary to reproduce the observed prevalence of quiescent galaxies as mass increases \citep[e.g.][]{Wetzel:2012lk}.  It is the same story when comparing RV to any of the other mass based models.  In Fig. \ref{fig:mstar_zform}, we show the the median redshift galaxies found it different mass haloes at $z=0$ reach a fraction, $f$, of their $z=0$ stellar mass.  Mass-based SHAM models generically predict much later growth compared to RV.  The one exception to this rule is the Y12 model, where at late times, massive galaxies lose stellar mass.  As a result, galaxies found in massive haloes reached $90\%$ of their $z=0$ mass relatively early.  Since we are not concerned with the evolution of massive galaxies, we do not investigate this aspect of Y12 any further in this paper.  Interestingly, even the RM model, where the SMHM relation does not evolve, results in galaxy growth much closer to the evolving models than with RV.  This is a consequence of the fact that haloes form their potential wells early, primarily adding mass in the outskirts at late times \citep{Bosch:2014cu}.  As a result, any model for galaxy formation based on $V_{\rm peak}$ will have to decouple galaxy growth from the growth of $V_{\rm peak}$ more than models based on $M_{\rm peak}$.

Finally, the effect of the dramatically different growth history in the RV model can be seen in the evolution of the stellar mass function.  In Fig. \ref{fig:stellar_mass_function_evolution} we show the predicted stellar mass function, $\phi(M_*,z)$, in the RV model and M13 (which was tuned to reproduce $\phi(M_*,z)$).  Here it is clear that the RV model over-produces galaxies, especially below the knee in the mass function.

@@

Given that the M13, Y12, and B13 models were fit to the stellar mass function at various redshifts and reproduce the cosmic star-formation density in the Universe (among other observables), it is difficult to imagine how to make a model like RV consistent with these same observational constraints.  While not explored here, it is likely that a more complicated $V_{\rm peak}$ based SHAM model could produce realistic galaxy growth histories, but this would require evolution in the SMVP relation.  Any evolution will require the SMVP relation be lower at high redshifts (as is the case for the SMHM relation in M13, Y12, and B13), reducing the stellar mass assigned to satellite galaxies in typical implementations.  This sort of modification would result in a decrease in the galaxy clustering signal on small scales.

Given this, it is not clear that there is a reason to prefer a $V_{\rm peak}$ based SHAM model over one based on $M_{\rm peak}$.  Apart from arguments related to fitting $w_p(r_p)$ and $\phi(M_*,z)$, there are other difficulties that must be overcome if $V_{\rm peak}$ is to be argued as more fundamental driving the stellar mass mapping into haloes.  For any given (sub-)halo, $V_{\rm peak}$ is generally set during major mergers \citep[1:5 or larger, ][]{Behroozi:2014bg}, and if stellar mass is tightly correlated with $V_{\rm peak}$, this would imply bursty star-formation closely tied to major mergers which is not favoured by observations \citep[e.g. star formation histories][]{Diemer:2017gj}.  On the other-hand, $M_{\rm peak}$ remains a theoretically attractive as a quantity that should be tightly coupled to stellar mass.  Peak halo mass is a good indicator of the amount of baryons that have been available to a galaxy to grow over its history {\bf [citations needed]}.

In our comparison between $V_{\rm peak}$-based SHAM and $V_{\rm peak}$-based SHAM it is clear that there is a tension between fitting galaxy clustering on small scales while simultaneously reproducing the build up of stellar mass in the Universe.  This tension presents a ``clustering crisis'' for SHAM.  Our exploration in the preceding sections has suggests multiple avenues SHAM models could be altered in order to alleviate this crisis.  We have fingered too few satellites as the culprit for the clustering deficiency in mass based SHAM models.  One solution is to simply increase the number of satellites in the mass-based models as many models have found to be necessary.  In \S \ref{sec:orphans}, we examine the plausibility of missing sub-haloes, and thus satellites, in our implementation of the mass-based SHAM models.  A significant population of missing sub-haloes would be an indication that so called ``oprhan'' galaxies play an important role in solving the clustering crisis.  Apart from orphans, we also consider two other physically motivated modifications to increase the satellite fraction--post-accretion satellite growth in \S \ref{sec:sat_growth} and assembly bias effects in \S \ref{sec:assem_bias}.

@@

Finally, we ask ``how many orphans are needed to increase the clustering signal sufficiently in the SHAM models based on $M_{\rm peak}$?''  To answer this question, we adjust $C_0$ in eq. (\ref{eq:sub_completeness}), while keeping the other parameters fixed, and fit to the galaxy clustering observations.  Lower values of $C_0$ imply an overall increased population of clone sub-haloes available to host orphans at all masses.  This correction assumes that sub-haloes are being artificially disrupted or merged with the host {\em at all masses}.  For each model, $C_0$ was adjusted in order to best fit the galaxy clustering signal in the intermediate mass bin $(10.0 \leq \log(M_*/h^{-2}M_{\odot}) <10.5)$.  The result of this exercise is shown in Fig. \ref{fig:wp_comparison_w_orphans} using the {\tt sub-profile} method to assign the positions and velocities.  We find that each model requires $C_{0} = [0.6,1.0, 0.4, 0.5, 0.5]$, for the RM, RV, M13, Y12, and B13 models respectively.  While the RV model does not require orphans, the mass-based SHAM models require that on average approximately {\em half of all satellites are orphans}.  In Fig. \ref{fig:orphan_clustering_2} we show how assigning orphans positions within their host affect the clustering predictions for the M13 model in the intermediate mass bin only.  As expected, using {\rm dm-profile} boosts the clustering signal at small scales relative to {\rm sub-profile}, but the effect is small compared to the dependence on $C_{0}$.  This trend is largely similar at the other mass considered and for the other SHAM models.

We make no attempt to fit for new parameters in the evolving SMHM models using our orphan model and clustering measurements, while the RM and RV models adjust automatically to the increased abundance of sub-haloes.  Adjusting the population of sub-haloes so drastically in the evolving models will have an effect on the parameter inference for the SMHM relation.  With this caveat in mind, the most noticeable failure of the mass-based models is the over-prediction of the small scale ($< 1 {\rm Mpc}$) clustering signal in the most massive stellar mass bin ($10.5\leq \log(M_*/h^{-2}M_{\odot})<11.0$).  Each of these models now produce too many massive satellites.  This problem could be reduced by altering the SMHM for massive satellites or reducing the number of massive orphans.  We leave a detailed study on the self-consistently including orphan galaxies in SHAM to future work.

@@

While we have shown that a large population of ``orphan'' galaxies would alleviate the clustering crisis in mass-based SHAM models, high resolution dark matter simulations do not seem to provide evidence of the requisite missing sub-halo population.  With this in mind, we explore alternative mechanisms to boost the galaxy clustering signal in SHAM models.  In this section, we consider whether continued star-formation in satellites after accretion can significantly boost clustering.  This is motivated by findings that suggest satellites continue to form stars and grow in stellar mass after accretion for between $\sim 2$ and $4$ Gyr \citep{Wetzel:2013dw}.  If satellites continue to grow after accretion, SHAM methods which use the SMHM relation at the time of accretion to assign stellar mass to satellites will under-estimate their stellar mass (e.g. M13 and B13).

@@

In this section we examine whether introducing assembly bias into the galaxy-halo mapping can increase the satellite fraction and galaxy clustering signal in mass-based SHAM models.  Specifically, we consider a model where $M_{*}$ is correlated with the formation history of the (sub-)halo it occupies such that early forming haloes host more massive galaxies than late forming haloes at {\em fixed} peak (sub-)halo mass.  While many measures of halo formation history have been employed, for this work we use the redshift a halo first achieves a fraction, $\rm f$, of its peak halo mass, $z_{\rm f}$ (see Appendix \ref{sec:halo_properties} for details on how $z_{\rm f}$ is calculated).

@@

To induce a correlation between $M_{*}$ and $z_{\rm f}$ we bin (sub-)haloes in small 0.1 dex $M_{\rm peak}$ bins.  We then rank order (sub-)haloes by $z_{\rm f}$ and galaxies by $M_{*}$, re-assigning the most massive galaxies to the earliest forming (sub-)haloes in the bin.  We use ${\rm f}=0.5$, the redshift where a (sub-)halo reaches half its peak mass.  Furthermore, we parametrize the strength of this correlation by the Spearman's rank order correlation coefficient, $\rho(M_{\rm vir})$.  We weaken the correlation by degrading the rank ordering as described in appendix \ref{appendix:ranks} in order to reduce the effect of assembly bias on $M_*$.
